%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Preface}

The primary body of this dissertation is based on a number of collaborative works that either have been published or are currently in the process of being reviewed. The development of all the results herein are based on the active collaboration between myself and my supervisor, Professor Michael P. Friedlander.

\begin{itemize}
    \item The material presented in \autoref{ch:Dual-Struc-Opt} is the result of additional collaboration with Dr. Halyun Jeong and Dr. Yifan Sun while they were postdocs at UBC. This work has been published in Foundations and Trends in Optimization, 2020; see \citet{fan2019alignment}. The main theoretical results of this work are developed by Yifan Sun and me, and the extension to the sum of sets is developed by me. Halyun Jeong provided many examples and useful suggestions. Michael P. Friedlander provided overall supervision, technical guidance, writing assistance and financial support. 
    \item The material presented in \autoref{ch:App-Sig-Demix} is the result of additional collaboration with Dr. Halyun Jeong and Dr. Babhru Joshi while they were postdocs at UBC. This work has been published in IEEE Transactions on Signal Processing, 2022; see \citet{fan2020polar}. The methodology, theoretical results, and numerical experiments are developed by me. Halyun Jeong and Babhru Joshi provided useful suggestions, improved the presentation of the work, and helped checking the proofs. Michael P. Friedlander provided overall supervision, technical guidance, writing assistance and financial support.
    \item The material presented in \autoref{ch:App-Primal-Retrieval} is the result of additional collaboration with Dr. Huang Fang while he was a PhD at UBC. This work has been submitted and made publicly available in preprint form; see \citet{fan2021safe}. The main theoretical results of this work are developed by Huang Fang and me, the case study on nuclear-norm regularized problem is developed by Huang Fang, and the numerical experiments are conducted by me. Michael P. Friedlander provided overall supervision, technical guidance, writing assistance and financial support.
    \item The material presented in \autoref{ch:App-AtomicOpt} is the documentation of a open-source package developed by myself and Michael P. Friedlander. The package is publicly available at \url{https://github.com/MPF-Optimization-Laboratory/AtomicOpt.jl}. 
    \item The material presented in \autoref{ch:Dual-Fed-Opt} is the result of additional collaboration with Dr. Huang Fang while he was a PhD at UBC. This work has been submitted and made publicly available in preprint form; see \citet{fan2022dual}. The methodology, theoretical results, and numerical experiments are developed by Huang Fang and me. Michael P. Friedlander provided overall supervision, technical guidance, writing assistance and financial support.
    \item The material presented in \autoref{ch:Val-HFL} is the result of additional collaboration with Dr. Huang Fang, Dr. Zirui Zhou, Professor Jian Pei, Dr. Changxin Liu, and Dr. Yong Zhang while I was doing an internship at Huawei Technology Canada. This work has been published in IEEE International Conference on Data Engineering, 2022; see \citet{fan2022improving}. The methodology, theoretical results, and numerical experiments are developed by me. Huang Fang and Zirui Zhou improved the presentation of the work, and helped checking the proofs. Changxin Liu and Yong Zhang participated in some discussions and provided some useful suggestions. Jian Pei and Michael P. Friedlander provided overall supervision, technical guidance and writing assistance. 
    \item The material presented in \autoref{ch:Val-VFL} is the result of additional collaboration with Dr. Huang Fang, Dr. Zirui Zhou, Professor Jian Pei, and Dr. Yong Zhang while I was doing an internship at Huawei Technology Canada. This work has been submitted and made publicly available in preprint form; see \citet{fan2022fair}. The methodology, theoretical results, and numerical experiments are developed by me. Huang Fang and Zirui Zhou improved the presentation of the work, and helped checking the proofs. Yong Zhang participated in some discussions and provided some useful suggestions. Jian Pei and Michael P. Friedlander provided overall supervision, technical guidance and writing assistance. 
\end{itemize}




 






